---
title: "Model selection and training"
author: "Fernando Campaña"
date: "1/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(dplyr)
```

## Cleaning the data

We parse the csv data:
```{r}
data <- read.csv("training.csv", na.strings = "?", sep = ",")  
test_file <- read.csv("test.csv", na.strings = "?", sep = ",")  
```

There are some columns with missing values (NA), so we have to filter these columns out. 

```{r}
data_filtered <- bind_cols(data[8:11], data[37:49], data["classe"])
test_filtered <- bind_cols(test_file[8:11], test_file[37:49])
```

## Cross validation
Then we need to split the data into a training and a testing set: 60% for training, 40% for testing.
```{r, cache=TRUE}
inTrain <- createDataPartition(y=data_filtered$classe, p=0.60,list=FALSE)
training <- data_filtered[inTrain,]
testing <- data_filtered[-inTrain,]
dim(training)
dim(testing)
```

## Training the model
We are planning on using Random Forest due to its ease of understanding and great accuracy. We proceed to train the model:
```{r, cache=TRUE}
modrf = train(classe ~ . , data= training, method="rf")
```

## Measuring error rates
We know that the out of sample error is going to be greater than the in sample error, due to overfitting.
```{r, cache=TRUE}
trainprediction <- predict(modrf, newdata = training)
table(trainprediction, training$classe)
correctTrain <-sum(trainprediction == training$classe)
correctTrain
correctTrain / length(trainprediction) * 100
testprediction <- predict(modrf, newdata = testing)
table(testprediction, testing$classe)
correctTest <- sum(testprediction == testing$classe)
correctTest
correctTest / length(testprediction) * 100
```
The Trainprediction table contains the info used to calculate the "in sample error", with 11776 correct answers and a 100% correct rate and 0% error rate.
While Testprediction is the "out of sample error", with 7522 correct answers and a 95.87% correct rate and a 5.13% error rate. These values confirm our prediction about error rates.

So we are going to use this model to predict the labels in the next assignment, by employing the following line. 
```{r, echo=FALSE}
predict(modrf, newdata = test_filtered)
```